Interactive Curvature Tensor Visualization on Digital Surfaces\thanks{This work has been mainly funded by \textsc{DigitalSnow} ANR-11-BS02-009, \textsc{KIDICO} ANR-2010-BLAN-0205 and \textsc{PRIMES Labex} ANR-11-LABX-0063/ ANR-11-IDEX-0007 research grants.}

### 2016-01-27 latex/article.tex
moved fig 7 to the bottom of the page 10
final

It has been shown that the radius of the ball depends on the geometry of the
underlying shape. In~\cite{DGCI2014}, a proposal was made for a parameter-free
estimation of the radius of the ball by analyzing the shape \wrt the local shape
geometry using maximal digital straight segments of the digital boundary.
In~\cite{SMI2015}, these estimators have been analyzed in scale-space (for a
range of radii) for a given digital shape. This allows to detect features of
the shape thanks to the behavior of estimators on singularities.

As a consequence, for all these integral invariant based approaches, we need to
consider different ball radius which could be time consuming when implemented on
CPU\@. We propose here a fully parallel implementation on GPU allowing us to
change the radius and thus update the estimated quantities in real-time.

@@

Representing a hierarchical structure on GPU is usually challenging since
such a data parallel component is unable to handle recursivity.

%since the GPU
%memory is limited and non consistent %\todo{fix coherent/aligned/coalesced?}
%memory access (\emph{e.g.} tree traversal using pointer dereferencing) may lead
%to huge payload from cache misses.

Efficient spatial tree encoding can be
achieved using pointerless structures such as linear quadtrees or octrees
%(seefor instance
 \textsc{Gargantini}~\cite{gargantini1982effective}. This structure
%consists in indexing each cell by a \emph{Morton code}: the code of children
indexes each cell by a \emph{Morton code}: the code of children
cells are defined by the code of the parent suffixed by two bits (in dimension
2, three bits in dimension 3) (see Figure
\ref{fig_quadtree_partitionning}-\emph{left}). A cell's code encodes its
position \wrt its parent cell and its complete path to the tree root.
%Hence, whatever their depth, all cell codes are simply stored in a linear vector on memory.
Hence, the tree is fully represented as a linear vector of its leaves.
%, whatever their depth.
Furthermore, cell operations such as subdivision, merging %, neighbors access
can be efficiently implemented using bitwise operations on the Morton
code. In the following, we use the GPU
friendly
implementation proposed by \textsc{Dupuy}
\textit{et al.}~\cite{dupuy2014quadtrees}.% that we extended to the third dimension.
%By using their approach, Morton codes
%and bitwise operations are optimized to match with GPU hardware requirements.

@@

It is also visible in Figure~\ref{fig:timings} that a hierarchical decomposition greatly reduces the curvature computation
time, especially with big radii.
%For the XYZ-Dragon with $R=20$, it needs almost
%4 seconds to compute the $l_0$ level, whereas using the hierarchical algorithm,
%it only needs 700ms.
However, due to the precomputation and the current hierarchical
structure, this algorithm is biased and creates an error (presented in
Table~\ref{tab:full-res-stat}) that needs to be considered.

\begin{table}
  \begin{center}
  \setlength{\tabcolsep}{0.15cm}
	\begin{tabular}{@{}llrrrrrr@{}}
      \toprule
        & & H & $l_4$ & $l_3$ & $l_2$ & $l_1$ & $l_0$\\
      \midrule
      \multirow{2}{2.2cm}{Number of texture fetches} & $R=8$ & 1468 & & 2 & 28 & 260 & 2104 \\
                                & $R=16$ & 5706 & 2 & 28 & 90 & 2120 & 17080\\\midrule

@@

      \multirow{2}{2.1cm}{$L_2$ error (\wrt $l_0$)} & $R=8$ & 3.51e-05 & & 2.67e-04 & 7.57e-05 & 2.02e-05 & 0\\
                               & $R=16$ & 4.43e-05 & 1.39e-04 & 4.16e-05 & 1.57e-05 & 2.07e-06 & 0\\
      \bottomrule
    \end{tabular}
  \end{center}
  %
  \caption{Comparison of number of texture fetches, $L_2$ and $L_\infty$ error
  obtained on ``OctaFlower'' with a digital domain of $130^3$ when computing mean
  curvature with two radii: 8 and 16, with hierarchical algorithm (H) and $l \ge
  l_0$ approximation algorithms. The object is triangulated at full resolution
  with a regular grid and contains 282,396 vertices.\label{tab:full-res-stat}}
  %
\end{table}

\begin{figure}[!htbp]
  \vspace{-0.35cm}
  \begin{center}
   {\includegraphics[width=2.7cm]{figs/octa_r10_l3_a}}
   {\includegraphics[width=2.7cm]{figs/octa_r10_l2_a}}
   {\includegraphics[width=2.7cm]{figs/octa_r10_l1_a}}
   {\includegraphics[width=2.7cm]{figs/octa_r10_l0_a}}
  \end{center}
  \vspace{-0.35cm}
  \caption{Illustration of mean curvature computation on ``OctaFlower'' (digital domain of $130^3$) using mipmap approximation with different levels: $l_3$, $l_2$, $l_1$ and $l_0$ (\ie no approximation).
  %
  %For all illustrations, we use directly normal vectors computed by algorithm discussed in Section~\ref{sec:curv-tens-estim}
  }\label{fig:refine}
  \vspace{-0.35cm}
\end{figure}

\begin{figure}[!htbp]
  \begin{center}
    \begin{tikzpicture}[scale=0.7]
      \begin{axis}[
          ybar stacked,
          ymode = log,
          ymax = 6000,
	  	  bar width=20pt,
	  	  nodes near coords,
	  	  every node near coord/.style={ font=\tiny, inner sep=0pt},
          enlargelimits=0.15,
          ylabel={\#Time (in ms)},
          symbolic x coords={Octa-R8, Octa-R16, Snow-R10, Snow-R20,
	      Dragon-R8, Dragon-R20},
          xtick=data,
          point meta=rawy,
          x tick label style={rotate=45,anchor=east},
        ]
        %Geometry Generation
        \addplot+[ybar, color=black, fill=cgeometry] plot coordinates {(Octa-R8,17.283000) (Octa-R16,18.458000)
          (Snow-R10,9.831000) (Snow-R20,9.653000)  (Dragon-R8,136.300000) (Dragon-R20,145.450000)};
        %l4
        \addplot+[ybar, color=black, fill=clquatre] plot coordinates {(Octa-R8,1) (Octa-R16,2.509000)
          (Snow-R10,6.085000) (Snow-R20,9.445000)  (Dragon-R8,3.389000) (Dragon-R20,20.228000)};
        %l3
        \addplot+[ybar, color=black, fill=cltrois] plot coordinates {(Octa-R8,2.527000) (Octa-R16,4.159000)
          (Snow-R10,10.629000) (Snow-R20,11.542000) (Dragon-R8,20.397000) (Dragon-R20,26.710000)};
        %l2
        \addplot+[ybar, color=black, fill=cldeux] plot coordinates {(Octa-R8,3.783000) (Octa-R16,9.006000)
          (Snow-R10,13.777000) (Snow-R20,37.700000) (Dragon-R8,29.967000) (Dragon-R20,92.435000)};
        %l1
        \addplot+[ybar, color=black, fill=clun] plot coordinates {(Octa-R8, 11.004000) (Octa-R16,38.687000)
          (Snow-R10,39.627000) (Snow-R20,217.954000) (Dragon-R8,59.840000) (Dragon-R20,506.549000)};
        %l0
        \addplot+[ybar, color=black, fill=clzero] plot coordinates {(Octa-R8, 37.273000) (Octa-R16,244.353000)
          (Snow-R10,236.194000) (Snow-R20,1655.680000) (Dragon-R8,322.674000) (Dragon-R20,3763.422000)};
        %\legend{\strut Geometry, \strut Curvature, \strut xx, \strut xx}
      \end{axis}
	  \end{tikzpicture}
%
    \begin{tikzpicture}[scale=0.7]
      \begin{axis}[
          ybar stacked,
          ymax = 6000,
          ymode = log,
	  	  bar width=20pt,
	  	  nodes near coords,
	  	  every node near coord/.style={ at={(0,0)}, font=\tiny, inner sep=0pt},
          enlargelimits=0.15,
          %ylabel={\#Time (in ms)},
          symbolic x coords={Octa-R8, Octa-R16, Snow-R10, Snow-R20,
	      Dragon-R8, Dragon-R20},
          xtick=data,
          point meta=rawy,
          x tick label style={rotate=45,anchor=east},
        ]
        %Geometry Generation
        \addplot+[ybar,color=black, fill=cgeometry] plot coordinates {(Octa-R8,17.283000) (Octa-R16,17.183000)
          (Snow-R10,9.383000) (Snow-R20,9.767000)  (Dragon-R8,127.347000) (Dragon-R20,127.166000)};
        %Hierarchical measurement
        \addplot+[ybar, color=black, fill=clzero] plot coordinates {(Octa-R8,18.095000) (Octa-R16,64.186000)
          (Snow-R10,96.747000) (Snow-R20,364.547000)  (Dragon-R8,147.662000) (Dragon-R20,787.459000)};
        %\legend{\strut Geometry, \strut Curvature, \strut xx, \strut xx}
      \end{axis}
     \end{tikzpicture}
     \end{center}
  %
  \vspace{-0.35cm}
  \caption{Timings in milliseconds (in logscale) obtained while visualizing an adaptive
  triangulation on three objects -- ``OctaFlower'' (digital domain of $130^3$), ``Snow microstructures'' ($233^3$) and ``XYZ-Dragon'' ($510^3$) --  by computing the
  curvature with a regular grid (\emph{left}) and with hierarchical
  algorithm (\emph{right}), with two different radii for each object.
  %
  In \emph{orange color}: time required to extract the triangulation.
  %
  In \emph{blue color}: time required to compute the curvature tensor at
  different levels: from $l_4$ (light blue) to $l_0$ (dark blue).
  %
  Timings are given using a NVIDIA GeForce GTX 850M GPU\@.
  }\label{fig:timings}
  \vspace{-0.35cm}
\end{figure}

 \begin{figure}[h!]
   \vspace{-0.35cm}
   \begin{center}
   	 {\includegraphics[width=0.4\textwidth]{figs/snow_r8}}
   	 {\includegraphics[width=0.4\textwidth]{figs/xyzrgb_dragon-510_R8_mean}}

     {\includegraphics[width=0.4\textwidth]{figs/snow_r8_min}}
	 {\includegraphics[width=0.4\textwidth]{figs/dragon_r8}}

     {\includegraphics[width=0.4\textwidth]{figs/snow_r8_max}}
	 {\includegraphics[width=0.4\textwidth]{figs/dragon_r8_max}}

     {\includegraphics[width=0.4\textwidth]{figs/snow_r8_normales}}
     {\includegraphics[width=0.4\textwidth]{figs/dragon_r8_normales}}

   \end{center}
   \caption{\emph{Left column}: Mean curvature, first and second
     principal directions and normal vector field estimation on ``Snow microstructures'' ($233^3$ and $R=8$).
  \emph{Right column:} Mean curvature, first and second principal
  directions and normal vector field estimation on ``XYZ-Dragon'' ($510^3$
  and $R=8$). Normal vectors are colored with a mapping of their
  component to RGB color space.}\label{fig:adaptive}
   \vspace{-0.35cm}
 \end{figure}

 Figure~\ref{fig:adaptive} shows curvature computation and exploration in
 real-time on large datasets: ``XYZ-Dragon'' (with digital domain of $512^3$) and
 ``Snow microstructures'' ($233^3$).

>>>>

It has been shown that the radius of the ball depends on the geometry of the
underlying shape. In~\cite{DGCI2014}, a proposal was made for a parameter-free
estimation of the radius of the ball by analyzing the shape \wrt the local shape
geometry using maximal digital straight segments of the digital boundary.
In~\cite{SMI2015}, these estimators have been analyzed in scale-space (for a
range of radii) for a given digital shape. This allows to detect features of
the shape thanks to the behavior of estimators on singularities.
As a consequence, for all these integral invariant based approaches, we need to
consider different ball radius which could be time consuming when implemented on
CPU. We propose here a fully parallel implementation on GPU allowing us to
change the radius and thus update the estimated quantities in real-time.

@@

Representing a hierarchical structure on GPU is usually challenging since
such a data parallel component is unable to handle recursivity.
%since the GPU
%memory is limited and non consistent %\todo{fix coherent/aligned/coalesced?}
%memory access (\emph{e.g.} tree traversal using pointer dereferencing) may lead
%to huge payload from cache misses.
Efficient spatial tree encoding can be
achieved using pointerless structures such as linear quadtrees or octrees
%(seefor instance
 \textsc{Gargantini}~\cite{gargantini1982effective}. This structure
%consists in indexing each cell by a \emph{Morton code}: the code of children
indexes each cell by a \emph{Morton code}: the code of children
cells are defined by the code of the parent suffixed by two bits (in dimension
2, three bits in dimension 3) (see Figure
\ref{fig_quadtree_partitionning}-\emph{left}). A cell's code encodes its
position \wrt its parent cell and its complete path to the tree root.
%Hence, whatever their depth, all cell codes are simply stored in a linear vector on memory.
Hence, the tree is fully represented as a linear vector of its leaves.
%, whatever their depth.
Furthermore, cell operations such as subdivision, merging %, neighbors access
can be efficiently implemented using bitwise operations on the Morton
code. In the following, we use the GPU
friendly
implementation proposed by \textsc{Dupuy}
\textit{et al.}~\cite{dupuy2014quadtrees}.% that we extended to the third dimension.
%By using their approach, Morton codes
%and bitwise operations are optimized to match with GPU hardware requirements.

@@

It is also visible in Figure~\ref{fig:timings} that a hierarchical decomposition greatly reduces the curvature computation
time, especially with big radii.
%For the XYZ-Dragon with $R=20$, it needs almost
%4 seconds to compute the $l_0$ level, whereas using the hierarchical algorithm,
%it only needs 700ms.
However, due to the precomputation and the current hierarchical
structure, this algorithm is biased and creates an error (presented in
Table~\ref{tab:full-res-stat}) that needs to be considered.
%
%
\begin{table}
  \begin{center}
  \setlength{\tabcolsep}{0.15cm}
	\begin{tabular}{@{}llrrrrrr@{}}
      \toprule
        & & H & $l_4$ & $l_3$ & $l_2$ & $l_1$ & $l_0$\\
      \midrule
      \multirow{2}{2.2cm}{Number of texture fetches} & $R=8$ & 1468 & & 2 & 28 & 260 & 2104 \\
                                & $R=16$ & 5706 & 2 & 28 & 90 & 2120 & 17080\\\midrule

@@

      \multirow{2}{2.1cm}{$L_2$ error (\wrt $l_0$)} & $R=8$ & 3.51e-05 & & 2.67e-04 & 7.57e-05 & 2.02e-05 & 0\\
                               & $R=16$ & 4.43e-05 & 1.39e-04 & 4.16e-05 & 1.57e-05 & 2.07e-06 & 0\\
      \bottomrule
    \end{tabular}
  \end{center}
  %
  \caption{Comparison of number of texture fetches, $L_2$ and $L_\infty$ error
  obtained on ``OctaFlower'' with a digital domain of $130^3$ when computing mean
  curvature with two radii: 8 and 16, with hierarchical algorithm (H) and $l \ge
  l_0$ approximation algorithms. The object is triangulated at full resolution
  with a regular grid and contains 282,396 vertices.\label{tab:full-res-stat}}
  %
\end{table}
%
\begin{figure}[!htbp]
  \vspace{-0.35cm}
  \begin{center}
   {\includegraphics[width=2.7cm]{figs/octa_r10_l3_a}}
   {\includegraphics[width=2.7cm]{figs/octa_r10_l2_a}}
   {\includegraphics[width=2.7cm]{figs/octa_r10_l1_a}}
   {\includegraphics[width=2.7cm]{figs/octa_r10_l0_a}}
  \end{center}
  \vspace{-0.35cm}
  \caption{Illustration of mean curvature computation on ``OctaFlower'' (digital domain of $130^3$) using mipmap approximation with different levels: $l_3$, $l_2$, $l_1$ and $l_0$ (\ie no approximation).
  %
  %For all illustrations, we use directly normal vectors computed by algorithm discussed in Section~\ref{sec:curv-tens-estim}
  }\label{fig:refine}
  \vspace{-0.35cm}
\end{figure}
%
\begin{figure}[!htbp]
  \begin{center}
    \begin{tikzpicture}[scale=0.7]
      \begin{axis}[
          ybar stacked,
          ymode = log,
          ymax = 6000,
	  	  bar width=20pt,
	  	  nodes near coords,
	  	  every node near coord/.style={ font=\tiny, inner sep=0pt},
          enlargelimits=0.15,
          ylabel={\#Time (in ms)},
          symbolic x coords={Octa-R8, Octa-R16, Snow-R10, Snow-R20,
	      Dragon-R8, Dragon-R20},
          xtick=data,
          point meta=rawy,
          x tick label style={rotate=45,anchor=east},
        ]
        %Geometry Generation
        \addplot+[ybar, color=black, fill=cgeometry] plot coordinates {(Octa-R8,17.283000) (Octa-R16,18.458000)
          (Snow-R10,9.831000) (Snow-R20,9.653000)  (Dragon-R8,136.300000) (Dragon-R20,145.450000)};
        %l4
        \addplot+[ybar, color=black, fill=clquatre] plot coordinates {(Octa-R8,1) (Octa-R16,2.509000)
          (Snow-R10,6.085000) (Snow-R20,9.445000)  (Dragon-R8,3.389000) (Dragon-R20,20.228000)};
        %l3
        \addplot+[ybar, color=black, fill=cltrois] plot coordinates {(Octa-R8,2.527000) (Octa-R16,4.159000)
          (Snow-R10,10.629000) (Snow-R20,11.542000) (Dragon-R8,20.397000) (Dragon-R20,26.710000)};
        %l2
        \addplot+[ybar, color=black, fill=cldeux] plot coordinates {(Octa-R8,3.783000) (Octa-R16,9.006000)
          (Snow-R10,13.777000) (Snow-R20,37.700000) (Dragon-R8,29.967000) (Dragon-R20,92.435000)};
        %l1
        \addplot+[ybar, color=black, fill=clun] plot coordinates {(Octa-R8, 11.004000) (Octa-R16,38.687000)
          (Snow-R10,39.627000) (Snow-R20,217.954000) (Dragon-R8,59.840000) (Dragon-R20,506.549000)};
        %l0
        \addplot+[ybar, color=black, fill=clzero] plot coordinates {(Octa-R8, 37.273000) (Octa-R16,244.353000)
          (Snow-R10,236.194000) (Snow-R20,1655.680000) (Dragon-R8,322.674000) (Dragon-R20,3763.422000)};
        %\legend{\strut Geometry, \strut Curvature, \strut xx, \strut xx}
      \end{axis}
	  \end{tikzpicture}
%
    \begin{tikzpicture}[scale=0.7]
      \begin{axis}[
          ybar stacked,
          ymax = 6000,
          ymode = log,
	  	  bar width=20pt,
	  	  nodes near coords,
	  	  every node near coord/.style={ at={(0,0)}, font=\tiny, inner sep=0pt},
          enlargelimits=0.15,
          %ylabel={\#Time (in ms)},
          symbolic x coords={Octa-R8, Octa-R16, Snow-R10, Snow-R20,
	      Dragon-R8, Dragon-R20},
          xtick=data,
          point meta=rawy,
          x tick label style={rotate=45,anchor=east},
        ]
        %Geometry Generation
        \addplot+[ybar,color=black, fill=cgeometry] plot coordinates {(Octa-R8,17.283000) (Octa-R16,17.183000)
          (Snow-R10,9.383000) (Snow-R20,9.767000)  (Dragon-R8,127.347000) (Dragon-R20,127.166000)};
        %Hierarchical measurement
        \addplot+[ybar, color=black, fill=clzero] plot coordinates {(Octa-R8,18.095000) (Octa-R16,64.186000)
          (Snow-R10,96.747000) (Snow-R20,364.547000)  (Dragon-R8,147.662000) (Dragon-R20,787.459000)};
        %\legend{\strut Geometry, \strut Curvature, \strut xx, \strut xx}
      \end{axis}
     \end{tikzpicture}
     \end{center}
  %
  \vspace{-0.35cm}
  \caption{Timings in milliseconds (in logscale) obtained while visualizing an adaptive
  triangulation on three objects -- ``OctaFlower'' (digital domain of $130^3$), ``Snow microstructures'' ($233^3$) and ``XYZ-Dragon'' ($510^3$) --  by computing the
  curvature with a regular grid (\emph{left}) and with hierarchical
  algorithm (\emph{right}), with two different radii for each object.
  %
  In \emph{orange color}: time required to extract the triangulation.
  %
  In \emph{blue color}: time required to compute the curvature tensor at
  different levels: from $l_4$ (light blue) to $l_0$ (dark blue).
  %
  Timings are given using a NVIDIA GeForce GTX 850M GPU\@.
  }\label{fig:timings}
  \vspace{-0.35cm}
\end{figure}
%
 \begin{figure}[h!]
   \vspace{-0.35cm}
   \begin{center}
   	 {\includegraphics[width=0.45\textwidth]{figs/snow_r8}}
   	 {\includegraphics[width=0.45\textwidth]{figs/xyzrgb_dragon-510_R8_mean}}

     {\includegraphics[width=0.45\textwidth]{figs/snow_r8_min}}
	 {\includegraphics[width=0.45\textwidth]{figs/dragon_r8}}

     {\includegraphics[width=0.45\textwidth]{figs/snow_r8_max}}
	 {\includegraphics[width=0.45\textwidth]{figs/dragon_r8_max}}

     {\includegraphics[width=0.45\textwidth]{figs/snow_r8_normales}}
     {\includegraphics[width=0.45\textwidth]{figs/dragon_r8_normales}}

   \end{center}
   \caption{\emph{Left column}: Mean curvature, first and second
     principal directions and normal vector field estimation on ``Snow microstructures'' ($233^3$ and $R=8$).
  \emph{Right column:} Mean curvature, first and second principal
  directions and normal vector field estimation on ``XYZ-Dragon'' ($510^3$
  and $R=8$). Normal vectors are colored with a mapping of their
  component to RGB color space.}\label{fig:adaptive}
   \vspace{-0.35cm}
 \end{figure}
 Figure~\ref{fig:adaptive} shows curvature computation and exploration in
 real-time on large datasets: ``XYZ-Dragon'' (with digital domain of $512^3$) and
 ``Snow microstructures'' ($233^3$).
